---
title: "Seurat Clustering R Notebook"
output: html_notebook
---
```{r}
library("fs")
library("remotes")
old_lib <- path_home_r("R/x86_64-pc-linux-gnu-library/old_lib/")
```

```{r}
library(dplyr)
library("Seurat")
library(patchwork)
library(ggplot2)
library(Connectome)
library(cowplot)
library(gridExtra)
```
```{r}
pd.data <- read.csv("~/PDAC/data/immune_only/NormalCells_RawDGE_210305.csv",sep=",", header=T,row.names=1, check.names=FALSE)
pd.meta <- read.csv("~/PDAC/data/immune_only/NormalCells_metadata_210305.csv",sep=",", header=T,row.names=1)
```
```{r}
genes <- row.names(pd.data)
#createSeuratObject for all cells
pd.full <- CreateSeuratObject(counts = pd.data, project = "pd",min.cells = 50, min.features = 200, is.expr=1, meta.data=pd.meta)
```
```{r}
Idents(pd.full) <- "cell.types.complex"
table(Idents(pd.full)) 
```
```{r}
pd.full <- NormalizeData(pd.full, normalization.method = "LogNormalize", scale.factor = 10000)
pd.full <- FindVariableFeatures(pd.full, selection.method = "vst", nfeatures = 1070) 
pd.full[["percent.mt"]] <- PercentageFeatureSet(object = pd.full, pattern = "^MT-")
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pd.full), 10)
```
```{r}
# Plot variable features with labels
plot1 <- VariableFeaturePlot(pd.full)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2
VlnPlot(pd.full, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.05)
```
```{r}
gene.names <- rownames(pd.full)
pd.full <- ScaleData(pd.full)
pd.full <- RunPCA(pd.full, verbose = FALSE) # default features are previously
```
#### Visualize PCA results
```{r}
print(pd.full[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(pd.full, dims = 1:2, reduction = "pca")
DimPlot(pd.full, reduction = "pca")
# Identify sig PCs
# direct visualization 
DimHeatmap(pd.full, dims = 1:15, cells = 500, balanced = TRUE)
```
#### find clusters
```{r}
pd.full <- FindNeighbors(pd.full, dims = 1:45) # only using significant PCs
pd.full <- FindClusters(pd.full)
pd.full <- RunUMAP(pd.full, dims = 1:45)
pd.full <- RunTSNE(pd.full, dims = 1:45, check_duplicates = FALSE)
```
#### plot clusters
```{r, fig.width=8, fig.height=6}
DimPlot(pd.full, label = TRUE) + NoLegend()
DimPlot(pd.full, group.by = "cell.types.complex", label = TRUE)
DimPlot(pd.full, label = TRUE, reduction = 'tsne') + NoLegend()
DimPlot(pd.full, label = TRUE, reduction = 'tsne', group.by = "cell.types.complex", repel = TRUE, cells = rownames(pd.meta)) + labs(title='Cell Type Visualization') + theme(plot.title = element_text(hjust = 0.5))
```
```{r, fig.width=8, fig.height=6}
a  = DimPlot(pd.full, label = TRUE, reduction = 'tsne', group.by = "cell.types.complex", repel = TRUE, cells = rownames(pd.meta)) + labs(title='Complex Cell Type Visualization') + theme(plot.title = element_text(hjust = 0.5))
ggsave('~/PDAC/visualization/TSNEplot_by_cell_types_complex.png', a, width=2.6, height=2, units="in", scale=3)

a = DimPlot(pd.full, label = TRUE, reduction = 'tsne', group.by = "cell.types", repel = TRUE, cells = rownames(pd.meta)) + labs(title='Cell Type Visualization') + theme(plot.title = element_text(hjust = 0.5))
ggsave('~/PDAC/visualization/TSNEplot_by_cell_types.png', a, width=2.6, height=2, units="in", scale=3)
```
```{r, fig.width=8, fig.height=6}
Idents(pd.full) <- "Tumor.class"
table(Idents(pd.full))
a = DimPlot(pd.full, label = TRUE, reduction = 'tsne', group.by = "Tumor.class", repel = TRUE, cells = rownames(pd.meta), label.size = 0) + labs(title='Grouped by Tumor Severity') + theme(plot.title = element_text(hjust = 0.5))
ggsave('~/PDAC/visualization/TSNEplot_by_tumor_severity.png', a, width=2.6, height=2, units="in", scale=3)
```
# Connectome
#### subset dataset
```{r}
pd.data <- read.csv("~/PDAC/data/immune_only/NormalCells_RawDGE_210305.csv",sep=",", header=T,row.names=1, check.names=FALSE)
pd.meta <- read.csv("~/PDAC/data/immune_only/NormalCells_metadata_210305.csv",sep=",", header=T,row.names=1)
pd.data <- pd.data[,rownames(pd.meta)]

pd <- CreateSeuratObject(counts = pd.data, project = "pd",min.cells = 50, min.features = 200, is.expr=1, meta.data=pd.meta)
```

```{r}
pd <- NormalizeData(pd)
connectome.genes <- union(Connectome::ncomms8866_human$Ligand.ApprovedSymbol,Connectome::ncomms8866_human$Receptor.ApprovedSymbol)
genes <- connectome.genes[connectome.genes %in% rownames(pd)]
pd <- ScaleData(pd,features = genes)
Idents(pd) <- 'cell.types.complex'
pd.con <- CreateConnectome(pd,species = 'human',min.cells.per.ident = 75,p.values = F,calculate.DOR = F)
```
#### filter edges
```{r}
p1 <- ggplot(pd.con, aes(x=ligand.scale)) + geom_density() + ggtitle('Ligand.scale')
p2 <- ggplot(pd.con, aes(x=recept.scale)) + geom_density() + ggtitle('Recept.scale')
p3 <- ggplot(pd.con, aes(x=percent.target)) + geom_density() + ggtitle('Percent.target')
p4 <- ggplot(pd.con, aes(x=percent.source)) + geom_density() + ggtitle('Percent.source')
a = plot_grid(p1,p2,p3,p4)
ggsave('~/PDAC/visualization/Connectome_distribution.png', a, width=3, height=2, units="in", scale=3)
```
```{r}
pd.con2 <- FilterConnectome(pd.con,min.pct = 0.1,min.z = 0.20,remove.na = T)
```
#### look into one receptor/ligand
```{r}
p1 <- NetworkPlot(pd.con2,features = 'VEGFA',min.pct = 0.1,weight.attribute = 'weight_sc',include.all.nodes = T)
p2 <- NetworkPlot(pd.con2,features = 'VEGFA',min.pct = 0.5,weight.attribute = 'weight_sc',include.all.nodes = T)
```
#### quick identification of top ligand producers and correlating receptor receivers
```{r}
a <- Centrality(pd.con2,
           modes.include = NULL,
           min.z = 1,
           min.pct = 0.35,
           weight.attribute = 'weight_sc',
           targets.include = c('CTL','T_Regs','DC','Fibroblast','Endothelial','SPP1_mono','C1QC_mono'),
           sources.include = c('CTL','T_Regs','DC','Fibroblast','SPP1_mono','C1QC_mono'),
           group.by = 'mode')
ggsave('~/PDAC/visualization/centrality_of_modes.png', a, width=3, height=2, units="in", scale=3)
```
#### quick identification of known ligand/receptor
```{r}
Centrality(pd.con2,
           modes.include = c('VEGF'),
           weight.attribute = 'weight_sc',
           min.z = 0,
           group.by = 'mechanism')
```
#### 2-cell interactome
```{r}
for (i in c('NaiveT','CTL','T_Regs','NK','DC','Fibroblasts','Endothelial','B_Cells')) {
p1 <- CellCellScatter(pd.con2,sources.include = 'SPP1_mono',targets.include = i,
                label.threshold = 3,
                weight.attribute = 'weight_sc',min.pct = 0.3,min.z = 0.8)
p1 <- p1 + xlim(0,NA) + ylim(0,NA)
print(p1)
}
for (i in c('NaiveT','CTL','T_Regs','NK','DC','Fibroblasts','Endothelial','B_Cells')) {
p1 <- CellCellScatter(pd.con2,targets.include = 'SPP1_mono',sources.include = i,
                label.threshold = 3,
                weight.attribute = 'weight_sc',min.pct = 0.3,min.z = 0.8)
p1 <- p1 + xlim(0,NA) + ylim(0,NA)
print(p1)
}
```

```{r}
for (i in c('NaiveT','CTL','T_Regs','NK','DC','Fibroblasts','Endothelial','B_Cells')) {
p1 <- CellCellScatter(pd.con2,sources.include = 'C1QC_mono',targets.include = i,
                label.threshold = 3,
                weight.attribute = 'weight_sc',min.pct = 0.3,min.z = 0.8)
p1 <- p1 + xlim(0,NA) + ylim(0,NA)
print(p1)
}
for (i in c('NaiveT','CTL','T_Regs','NK','DC','Fibroblasts','Endothelial','B_Cells')) {
p1 <- CellCellScatter(pd.con2,targets.include = 'C1QC_mono',sources.include = i,
                label.threshold = 3,
                weight.attribute = 'weight_sc',min.pct = 0.3,min.z = 0.8)
p1 <- p1 + xlim(0,NA) + ylim(0,NA)
print(p1)
}
```

```{r}
png('~/PDAC/visualization/CircosPlot/TAM_asSources.png', width=8, height=5, units = 'in', res = 300)
CircosPlot(pd.con,sources.include = c('C1QC_mono','SPP1_mono','FCN1_mono') ,lab.cex = 0.6, min.z = 1,targets.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial', 'B_Cells'))
dev.off()
```

```{r}
test <- data.frame(pd.con2 %>% group_by(vector) %>% top_n(5,weight_sc))
png('~/PDAC/visualization/CircosPlot/C1QC_asTarget.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,targets.include = 'C1QC_mono',lab.cex = 0.6, min.z = 1,
           sources.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),
           title = "C1QC+ Macrophages as Target")
dev.off()
png('~/PDAC/visualization/CircosPlot/C1QC_asSource.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,targets.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),lab.cex = 0.6, min.z = 1,
           sources.include = 'C1QC_mono', edge.color.by.source = FALSE,
           title = "C1QC+ Macrophages as Source")
dev.off()
png('~/PDAC/visualization/CircosPlot/SPP1_asTarget.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,targets.include = 'SPP1_mono',lab.cex = 0.6, min.z = 1,
           sources.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),
           title = "SPP1+ Macrophages as Target")
dev.off()
png('~/PDAC/visualization/CircosPlot/SPP1_asSource.png', width=7, height=5, units = 'in', res = 300)
a = CircosPlot(test,targets.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),lab.cex = 0.6, min.z = 1,
           sources.include = 'SPP1_mono', edge.color.by.source = FALSE,
           title = "SPP1+ Macrophages as Source")
dev.off()

test <- data.frame(pd.con2 %>% group_by(vector) %>% top_n(5,weight_sc))
png('~/PDAC/visualization/CircosPlot/FCN1_asTarget.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,targets.include = 'FCN1_mono',lab.cex = 0.6, min.z = 1,
           sources.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),
           title = "FCN1+ Macrophages as Target")
dev.off()
png('~/PDAC/visualization/CircosPlot/FCN1_asSource.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,targets.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),lab.cex = 0.6, min.z = 1,
           sources.include = 'FCN1_mono', edge.color.by.source = FALSE,
           title = "FCN1+ Macrophages as Source")
dev.off()

png('~/PDAC/visualization/CircosPlot/TAM_asTarget.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,sources.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),lab.cex = 0.6, min.z = 1,
            targets.include = c('FCN1_mono','SPP1_mono','C1QC_mono'), edge.color.by.source = FALSE,
            title = "Macrophages as Target")
dev.off()
png('~/PDAC/visualization/CircosPlot/TAM_asSource.png', width=7, height=5, units = 'in', res = 300)
CircosPlot(test,targets.include = c('NaiveT','CTL','T_Regs','NK','DC','Fibroblast','Endothelial'),lab.cex = 0.6, min.z = 1,
            sources.include = c('FCN1_mono','SPP1_mono','C1QC_mono'), edge.color.by.source = FALSE,
            title = "Macrophages as Source")
dev.off()
```
#### check for ligand/receptor expression 
```{r}
VlnPlot(pd.full, idents = c("SPP1_mono","C1QC_mono"), features = "ITGAM")
VlnPlot(pd.full, split.by = "Tumor.class", features = "TGF", group.by = "cell.types")
```

### differential analysis between basal and classical subtypes
```{r}
# Split the object by condition:
cb.list <- SplitObject(pd, split.by = 'Tumor.class')
# Normalize, Scale, and create Connectome:
cb.con.list <- list()
for (i in 1:length(cb.list)){
  cb.list[[i]] <- NormalizeData(cb.list[[i]])
  cb.list[[i]] <- ScaleData(cb.list[[i]],features = rownames(cb.list[[i]]))
  Idents(cb.list[[i]]) <- 'cell.types'
  cb.con.list[[i]] <- CreateConnectome(cb.list[[i]],species = 'human',p.values = F)
}
names(cb.con.list) <- names(cb.list)
```
#### make differential connectome
```{r}
diff <- DifferentialConnectome(cb.con.list[[1]],cb.con.list[[2]])
```
#### find significant
```{r}
# Stash idents and make new identities which identify each as stimulated vs. control
Idents(pd) <- "cell.types"
celltypes <- as.character(unique(Idents(pd)))
celltypes.classical <- paste(celltypes, 'Classical', sep = '_')
celltypes.basal <- paste(celltypes, 'Basal', sep = '_')
Idents(pd) <- "cell.types"
pd$celltype.condition <- paste(Idents(pd), pd$Tumor.class, sep = "_")
#pd$celltype <- Idents(pd)
Idents(pd) <- "celltype.condition"

# Identify which ligands and receptors, for which cell populations, have an adjusted p-value < 0.05 based on a Wilcoxon rank test
diff.p <- data.frame()
for (i in 1:length(celltypes)){
  temp <- FindMarkers(pd,
                      ident.1 = celltypes.basal[i],
                      ident.2 = celltypes.classical[i],
                      verbose = FALSE,
                      features = genes,
                      min.pct = 0.25,
                      logfc.threshold = 0.25)
  temp2 <- subset(temp, p_val_adj < 0.05)
  if (nrow(temp2)>0){
  temp3 <- data.frame(genes = rownames(temp2),cells = celltypes[i])
  diff.p <- rbind(diff.p, temp3)
  }
}
diff.p$cell.gene <- paste(diff.p$cells,diff.p$genes,sep = '.')

# Filter differential connectome to only include significantly perturbed edges
diff$source.ligand <- paste(diff$source,diff$ligand,sep = '.')
diff$target.receptor <- paste(diff$target,diff$receptor,sep = '.')
diff.sub <- subset(diff,source.ligand %in% diff.p$cell.gene & target.receptor %in% diff.p$cell.gene)
```
```{r}
png('~/PDAC/visualization/differential_scoring.png', width=7, height=5, units = 'in', res = 300)
a <- DifferentialScoringPlot(diff.sub,min.score = 0.5,min.pct = 0.1,infinity.to.max = T)
ggsave('~/PDAC/visualization/differential_scoring.png', a, width=3, height=2, units="in", scale=3)
```
```{r}
png('~/PDAC/visualization/differential_circos.png', width=7, height=5, units="in", res=300)
a <- CircosDiff(diff.sub,min.score = 0.5,min.pct = 0.1,lab.cex = 0.4)
dev.off()
```

```{r}
CircosDiff(diff.sub,min.score = 0.5,min.pct = 0.1,lab.cex = 0.4,
           sources.include = c('Macrophage'), targets.include = c('T_Cells','Fibroblasts','DC'), edge.color.by.source = FALSE)
CircosDiff(diff.sub,min.score = 0.2,min.pct = 0.1,lab.cex = 0.4,
           targets.include = c('Macrophage'),sources.include = c('T_Cells','Fibroblasts','DC'))
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
